---
title: "Agriculture"
author: "Elise Gertsch and Spencer Wadsworth"
date: "April 4, 2019"
output: word_document
---

```{r}
agr <- read.table("https://mheaton.byu.edu/Courses/Stat469/Topics/3%20-%20SpatialCorrelation/1%20-%20PointReference/HWCaseStudy/Data/WaterHoldingCapacity.txt", header = TRUE)

library(ggplot2)
library(gridExtra)
library(tidyverse)
library(car)
library(geoR)
source("https://raw.githubusercontent.com/MJHeaton/glstools/master/stdres.gls.R")
source("https://raw.githubusercontent.com/MJHeaton/glstools/master/predictgls.R")
library(multcomp)
library(nlme)

pred <- agr %>% filter(is.na(WHC)) 
obs <- agr %>% filter(!is.na(WHC))
```

1. Create exploratory plots of the data by looking at the relationship between WHC (the response variable) and Yield and EC. Comment on any general relationships you see from the data.
```{r}
ggplot(data = obs, aes(x = EC, y = WHC)) + geom_point() + xlab("Electroconductivity at the location") + ylab("Water Holding Capacity at the location")
#The data appears to be linear with WHC increasing as EC increases

ggplot(data = obs, aes(x = Yield, y = WHC)) + geom_point() + xlab("Electroconductivity at the location") + ylab("Water Holding Capacity at the location")
#The data appears to be linear with WHC increasing as Yield increases

#These plots are just for our benefit and curiosity
ggplot(data = obs, aes(x = Lon, y = Lat, fill = Yield)) + geom_raster() + scale_fill_distiller(palette = "Spectral", na.value = NA)

ggplot(data = obs, aes(x = Lon, y = Lat, fill = EC)) + geom_raster() + scale_fill_distiller(palette = "Spectral", na.value = NA)
```

2. Fit an independent MLR model with a linear effect between Yield, EC and the response variable WHC. Explore the residuals to see if there is evidence of spatial correlation by mapping the residuals and plotting the variogram of the residuals.
```{r}
agr_lm <- lm(WHC ~ Yield*EC, data = obs)
summary(agr_lm)

ggplot(data = obs, aes(x = Lon, y = Lat, fill = resid(agr_lm))) + geom_raster() + scale_fill_distiller(palette = "Spectral", na.value = NA)
#certain sections of the map have related colors. This suggests correlation

vario <- variog(coords=obs[,1:2], data=resid(agr_lm))
plot(vario)
#variogram appears to be increasing, this suggesting correlation in the residuals
```

3. To determine an appropriate correlation structure to use, fit a spatial model using exponential, spherical and Gaussian correlation functions with a nugget effect (don't forget to filter out the missing observations). Compare the model fits using AIC and use the best fit model for the remainder of the analysis.
```{r}
exp.mod <- gls(model = WHC ~ Yield*EC, data = obs, correlation = corExp(form=~Lon + Lat, nugget = TRUE), method = "ML")
sph.mod <- gls(model = WHC ~ Yield*EC, data = obs, correlation = corSpher(form=~Lon + Lat, nugget = TRUE), method = "ML")
gauss.mod <- gls(model = WHC ~ Yield*EC, data = obs, correlation = corGaus(form=~Lon + Lat, nugget = TRUE), method = "ML")

AIC(exp.mod)
AIC(sph.mod)
AIC(gauss.mod)
#the exponential model yields the smallest AIC value. We will use that for the rest of the analysis

#just to see the constrained variables
coef(exp.mod$modelStruct$corStruct, unconstrained = FALSE)
```

4. Write out your model for analyzing the agriculture data in terms of parameters. Explain and interpret any parameters associated with the model.
```{r}
# y ~ N(X*β, σ^2*R)
# y is the response variable, in this case WHC
# X is the matrix with the various parameters plus an intercept. The parameters include EC, Yield, and the interaction between the two
# β is the vector of coefficients for each of the variables.
# σ^2*R is a covariance matrix. 
#R is the correlation matrix that has the correlation values between different the various coordinates. This results in 1s in the diagonal. In our case the correlation follows the exponential model so                     ρ(si, sj) = exp{-||si - sj||/ϕ}
```

5. Fit your spatial MLR model and validate any assumptions you made to fit the model.
```{r}
# 1. linearity
avPlots(lm(WHC ~ EC*Yield, data = obs))
# the AV plots show that our data is linear

# 2. independence
d.resids <- stdres.gls(exp.mod)
coords <- as.matrix(obs[,1:2], ncol = 2)
plot(variog(coords = coords, data = d.resids))
# the variogram of decorrelated residuals is a fairly straight line, so we can say that it shows independence

ggplot(data = obs, aes(x = Lon, y = Lat, fill = stdres.gls(exp.mod))) + geom_raster() + scale_fill_distiller(palette = "Spectral", na.value = NA)
#appears at least slightly more random than the last one

# 3. normality of residuals
ggplot(data = obs, aes(x = d.resids)) + geom_histogram(binwidth = .8) + xlab("Standardized Residuals") + ylab("Frequency")
ks.test(d.resids, 'pnorm')
# the histogram is slightly skewed left
#the high p-value on the ks test helps us feel more comfortable about the normality

# 4. equal variance
ggplot(data = obs, aes(x = fitted(exp.mod), y = d.resids)) + geom_point() + geom_abline(intercept = 0, slope = 0) + labs(x = "Fitted Values", y = "Residuals") + ggtitle("Checking Equal Variance")
# we can say that the condition of equal variance is met because the points follow a distribution for each of the five levels of Surface

```


6. Carry out a hypothesis test that locations with higher yield had higher WHC (which would make sense because more water would be available for the plant to use). Include a confidence interval for the effect of Yield on WHC and interpret this interval.
```{r}
summary(exp.mod)
#the p-value associated with an increasing WHC with yield is 0.0236. this is significant and suggests that locations with higher yield have higher WHC

intervals(exp.mod)$coef
#the confidence interval for the Yield coefficient is 0.017, 0.23 which does not contain 0, so WHC apparently increases with Yield
```

7. Predict WHC at all the locations where WHC is missing. Provide a plot of your predictions.
```{r}
preds <- predictgls(exp.mod, newdframe = pred)

ggplot(data = preds, aes(x = Lon, y = Lat, fill = Prediction)) + geom_raster() + scale_fill_distiller(palette = "Spectral", na.value = NA)
```


8. (Extra Credit) Add a google map of the satellite terrain to your prediction map in #7.
