---
title: "cardiovascular health"
author: "Elise Gertsch"
date: "March 18, 2019"
output: word_document
---

```{r message=FALSE, warning=FALSE}
heart <- read.table("https://mheaton.byu.edu/Courses/Stat469/Topics/2%20-%20TemporalCorrelation/2%20-%20Longitudinal/HWCaseStudy/Data/Tachycardia.txt", header = TRUE)

library(ggplot2)
library(nlme)
library(car)
source("https://raw.githubusercontent.com/MJHeaton/glstools/master/stdres.gls.R")
library(multcomp)

heart$SEX <- as.factor(heart$SEX)
heart$PERIOD <- as.factor(heart$PERIOD)
```

1. Create exploratory plots of looking at the relationship between HEARTRTE (the response variable) and some of the explanatory variables. Comment on any general relationships you see from the data.
```{r}
ggplot(data = heart, aes(x = PERIOD, y = HEARTRTE, colour = SEX)) + geom_boxplot() + scale_color_hue(labels = c("Male", "Female")) + ylab("Heart Rate") + ggtitle("Heart Rate by Period and Sex")
# for periods 1 and 3, women have a slightly higher median heart rate

ggplot(data = heart, aes(x = HEARTRTE)) + geom_density(fill = "darkblue")
```

2. Fit an independent MLR model with a linear effect of all variables except RANDID and PERIOD. Explore the residuals to see if there is evidence of correlation within patients from period to period (visit to visit).
```{r}
model <- lm(log(HEARTRTE) ~ .-RANDID - PERIOD, data = heart)

# beta-hat coefficients
coef(model)

# histogram of the residuals
ggplot(data = heart, aes(x = model$residuals)) + geom_histogram()
```

3. To determine an appropriate correlation structure to use, fit a longitudinal MLR model with an AR1, MA1 and general symmetric correlation matrix within each patient but independent across patients. Compare the model fits using AIC (which can be extracted from a gls() object using AIC()).
```{r}
mod_long1 <- gls(model = log(HEARTRTE) ~ .-RANDID - PERIOD, data = heart, corSymm(form =~ 1 | RANDID), method = "ML")
mod_long2 <- gls(model = log(HEARTRTE) ~ .-RANDID - PERIOD, data = heart, corAR1(form =~ 1 | RANDID), method = "ML")
mod_long3 <- gls(model = log(HEARTRTE) ~ .-RANDID - PERIOD, data = heart, corARMA(form =~ 1 | RANDID, p = 0, q = 1), method = "ML") #ARIMA with p = 0 and q = 1

# AIC
AIC(model)
AIC(mod_long1) # lowest AIC
AIC(mod_long2)
AIC(mod_long3)

mod_long <- mod_long1
```

4. Write out your model for analyzing the Tachycardia data in terms of parameters. Explain and interpret any parameters associated with the model.
```{r}

```

5. Fit your longitudinal model and validate any assumptions you made to fit the model.
```{r}
# linearity 
# can use our independent model because this isn't impacted by the longitudinal nature of the data
avPlots(model)

# independence
iresids <- stdres.gls(mod_long)
iresid <- matrix(stdres.gls(mod_long), byrow = TRUE, ncol = 3)
cor(iresid)
# the correlations are close to 0, so we can say that our assumption of independence is met

# normality of residuals
hist(iresids, xlab = "Decorrelated Residuals", main = "Histogram of Decorrelated Residuals")
# approximately normal, so we will say that normality of the residuals is met

# equal variance
ggplot(data = heart, aes(x = fitted(mod_long), y = iresids)) + geom_point() + geom_hline(yintercept = 0, col = "red")
# doesn't look very normally varied, but it is evenly distributed above and below 0 

```

