---
title: "cardiovascular health"
author: "Elise Gertsch"
date: "March 18, 2019"
output: word_document
---

```{r message=FALSE, warning=FALSE}
heart <- read.table("https://mheaton.byu.edu/Courses/Stat469/Topics/2%20-%20TemporalCorrelation/2%20-%20Longitudinal/HWCaseStudy/Data/Tachycardia.txt", header = TRUE)

library(ggplot2)
library(nlme)
heart$SEX <- as.factor(heart$SEX)
heart$PERIOD <- as.factor(heart$PERIOD)
```

1. Create exploratory plots of looking at the relationship between HEARTRTE (the response variable) and some of the explanatory variables. Comment on any general relationships you see from the data.
```{r}
ggplot(data = heart, aes(x = PERIOD, y = HEARTRTE, colour = SEX)) + geom_boxplot() + scale_color_hue(labels = c("Male", "Female")) + ylab("Heart Rate") + ggtitle("Heart Rate by Period and Sex")
# for periods 1 and 3, women have a slightly higher median heart rate

ggplot(data = heart, aes(x = HEARTRTE)) + geom_density(fill = "darkblue")
```

2. Fit an independent MLR model with a linear effect of all variables except RANDID and PERIOD. Explore the residuals to see if there is evidence of correlation within patients from period to period (visit to visit).
```{r}
model <- lm(HEARTRTE ~ .-RANDID - PERIOD, data = heart)

# beta-hat coefficients
coef(model)

# histogram of the residuals
ggplot(data = heart, aes(x = model$residuals)) + geom_histogram()
```

3. To determine an appropriate correlation structure to use, fit a longitudinal MLR model with an AR1, MA1 and general symmetric correlation matrix within each patient but independent across patients. Compare the model fits using AIC (which can be extracted from a gls() object using AIC()).
```{r}
mod_long <- gls(model = )
```

4. Write out your model for analyzing the Tachycardia data in terms of parameters. Explain and interpret any parameters associated with the model.
```{r}

```

5. Fit your longitudinal model and validate any assumptions you made to fit the model.
```{r}

```

